The importance of measurement uncertainty in ecological management

### 2018-08-16 manuscripts/manuscript.tex
WIP

  \begin{abstract}
  Ecological management and decision-making typically focus on uncertainty
  about the future, but surprisingly little is known about how to account
  for uncertainty of the present: that is, the realities of having only
  partial or imperfect measurements. Our two primary paradigms for
  handling decisions under uncertainty -- the precautionary principle and
  optimal control -- have so far given entirely contradictory results.
  This paradox is best illustrated in the example of fisheries management,
  where many ideas that guide thinking about ecological decision making
  were first developed. We find that simplistic optimal control approaches
  have repeatedly concluded that a manager should increase catch quotas
  when faced with greater uncertainty about the fish biomass. Current best
  practices take a more precautionary approach, decreasing catch quotas by
  a fixed amount to account for uncertainty. Using comparisons to both
  simulated and historical catch data, we find that neither approach is
  sufficient to avoid stock collapses under moderate observational
  uncertainty. Using partially observed Markov decision process (POMDP)
  methods, we demonstrate how this paradox arises from flaws in the
  standard theory, which contributes to over-exploitation of fisheries and
  increased probability of economic and ecological collapse. In contrast,
  we find POMDP-based management avoids such over-exploitation while also
  generating higher economic value. These results have significant
  implications for how we handle uncertainty in both fisheries and
  ecological management more generally.
  \end{abstract}
   \begin{keyword} Measurement uncertainty \sep POMDP \sep Decision theory \sep Fisheries \sep Conservation \sep \end{keyword}
 \end{frontmatter}

Imperfect information is ubiquitous in ecological management and
conservation decision making. While the pressing concerns of global
change have put the spotlight on \emph{forecasting}, i.e.~the
uncertainty of the future (e.g. Petchey et al. 2015), management
decisions must also contend with uncertainty of the \emph{present}: How
many fish are in the sea today? What regions harbor the most fragile
biodiversity? How have invasive species transformed native communities?
A recognition of the importance of uncertainty as well as sophisticated
methods and tools to do so has become increasingly widespread in the
context of estimating ecological models from available data, thanks in
large part to the rise of approaches such as Hierarchical Bayesian
Modeling, along with convenient software tools for implementing these
methods. In sharp contrast to that shift, the treatment of uncertainty
in decisions -- how the uncertainty in models should be reflected in
scientific recommendations, remains far less developed, more
qualitative, and under-served by available tools. In this paper we
illustrate how uncertainty in measurement has a dramatic negative impact
on outcomes under a variety of these existing approaches, which can be
avoided through the use of more formal approaches that have since been
developed in other fields that can likewise be implemented in software
tools. This insight both resolves a long-standing paradox in fisheries
management and has important implications for how ecosystem management
deals with uncertainty more generally.

Approaches for decision making under uncertainty in ecological systems
can be divided into two camps: approaches based in ``optimal control''
and usually favored by natural resource economists, and approaches based
in heuristic methods such as scenario planning, resilience thinking, and
precautionary rules of thumb, more commonly found in both ecological
literature and actual practice (Fischer et al. 2009; Polasky et al.
2011). While researchers have for some time recognized the need to unify
the transparent and quantitative algorithmic approach of optimal control
with the greater complexity and uncertainty of real ecosystems that is
acknowledged by heuristic methods, computational barriers to doing so
have hither-to stymied this progress. Here, we illustrate how the
limitations of these approaches has been manifested in the example of
fisheries management, and present a new approach that can combine the
reality of measurement uncertainty and the rigor of optimization to
resolve a long-standing paradox and suggest a more robust approach to
management.

Fisheries conservation and management has long been both a crucible and
proving ground for the theory of ecological management more generally,
including topics such as adaptive management (Walters and Hilborn 1978),
ecosystem-based management (Levin and Lubchenco 2008) and resilience
thinking (Holling 1973; May 1977), while also giving rise to the
sub-discipline of resource economics (Gordon and Press 1954; Schaefer
1954; Beverton and Holt 1957), thanks to its global relevance, long
history, and readily available data. Consequently, understanding the
impact of measurement uncertainty in this rich and well-studied context
will have implications for both resource management and conservation
efforts in other domains, many of which draw on concepts established or
tested in fisheries ecosystems. In particular, the approach here
illustrates the limitations of both over-simple optimal control
approaches and rule-of-thumb precautionary approaches, and offers
instead a more general strategy for dealing with measurement uncertainty
adaptively.

We compare the results of managing simulated fish stocks under the
existing optimal control approach, a heuristic approach that more
closely resembles today's management practices, and a newly proposed
approach that provides a more rigorous treatment of uncertainty.

\section{The Decision Problem}\label{the-decision-problem}

We consider the management problem of setting catch quotas for a marine
fishery in the face of imperfect information about the current stock
size and uncertainty about future recruitment.

We seek to determine the sequence of actions \(h_t\) for
\(t \in [1, ... \infty]\) that maximize the net present value
(discounted sum of all future profits) of the fishery. We will denote
the discount rate \(\gamma\). For simplicity we will once again follow
classic theory and assume a fixed price for fish, (equivalently,
measuring our value in units of discounted fish rather than discounted
dollars), though again this assumption is easy to relax. Each year, the
manager also obtains an estimate \(y_t\) of the true \(x_t\) stock size
subject to some measurement uncertainty \(\xi_t\)

\[y_t = \xi_t x_t\]

@@

\[\max_{ \lbrace h_t \rbrace } \mathbb{E} \left \{ \sum_{t=0}^{\infty} \gamma^t \cdot h_t \right \}\]

@@

\subsection{Population model}\label{population-model}

@@

\[x_{t+1} = \varepsilon_t (x_t-h_t) r  \left(1 - \frac{(x_t-h_t)}{K}\right) \]

where \(x_t\) is the current stock size, \(h_t\) the harvest chosen that
year, with parameters \(r\) giving the individual growth rate, and \(K\)
the carrying capacity and \(\varepsilon\) representing stochastic
recruitment in a variable environment. We have written this in terms of
\(x_t - h_t\) to underscore that we assume the convention of observe,
harvest, recruit. We will assume for simplicity
\(\varepsilon \sim \mathcal{N}(1, \sigma_g)\). This and similar models
are widely used in large scale analyses across diverse stocks (e.g.
(Costello et al. 2016; Britten et al. 2017)), and form the basis for
much of bioeconomic theory (Clark 1990). As such, it will be easier to
compare our results against intuition and classic theory, and avoid the
possibility of differences arsing only because of some particular subtle
assumption hidden in a more complex model.

It is important to remember that both existing methods we discuss and
the POMDP approach proposed here can or are already applied to more
complex models as well, including those with age or stage structure.
Likewise, while we consider the model and it's parameters as ``given,''
it should be understood to mean that this in practice comes as the
result of some model choice and statistical estimation from historical
data, frequently accounting for uncertainty in both the process and the
measurements (e.g. Dichmont et al. 2016). As we shall see here, having
included uncertainty in the model \emph{estimation} in no way excuses us
from having to also deal once again with that uncertainty in the
decision process.

\section{Current Theory and Practice}\label{current-theory-and-practice}

@@

\subsection{Uncertainty in Practice: MSY \&
TAC}\label{uncertainty-in-practice-msy-tac}

Maximum Sustainable Yield (MSY) remains the basis of international law
(including the UN, IWC, IATTC, ICCAT, ICNAF; Mace (2001)) and a familiar
standard of management in terrestrial ecosystems as well as aquatic
(Clark 1990). Critics have for some time observed the limitations of
harvesting at MSY in face of uncertainty in stock sizes and population
dynamics (Larkin 1977; Botsford et al. 1997). Many US fisheries reflect
this uncertainty through a series of adjustments that effectively reduce
the target fishing mortality level to reflect this uncertainty.
Typically, a stock assessment model provides a best estimate of
\(B_{MSY}\) and corresponding mortality \(F_{MSY}\) is used to define
the stock Over-fishing Limit, (OFL). Based on this, a somewhat lower
level is set as the Allowable Biological Catch (ABC), reflecting
uncertainty in the stock assessment. To reflect possible uncertainty
between reported and actual catch, the ABC is reduced somewhat further
to define the Total Allowable Catch, (TAC), which forms the basic unit
of management for many such fisheries. To reflect this process, our
analysis will also consider policies in which the harvest quota is set
at 80\% of the level expected under the MSY policy:
\(H_{TAC}(t) = 0.8 \cdot F_{MSY} \cdot B_t\), for a biomass estimated at
\(B(t)\) in year \(t\). To distinguish this approach from MSY, we will
refer to this approach as a TAC policy. This more closely represents a
heuristic (e.g. Hilborn 2010) or resilience-based approach (sensu
Fischer et al. 2009) than an optimization-based policy. Importantly,
this approach shares the fundamentally stationary assumptions of an MSY
policy by defining a constant mortality rather than a dynamic policy.
Thus, despite being more cautious overall and generating lower economic
yield than a constant escapement policy, TAC policies continue to
harvest at non-zero rates even if a stock falls below \(B_{MSY}\), while
the constant escapement policy does not.

\subsection{Optimal management}\label{optimal-management}

So far, attempts to provide a more formal basis for managing uncertainty
than the heuristic adjustment of catch limits described above have
largely foundered in a series of paradoxes. Without uncertainty, the
theoretical policies are quite intuitive: The result derived by Clark
(1973) for an optimal dynamic strategy to replace the equilibrium
solution of MSY can be summarized as: If the stock is most productive at
\(B_{MSY}\), then obtain \(B_{MSY}\) as quickly as possible. Thus, at
any level below \(B_{MSY}\), the economically optimal thing to do is to
completely shut down the fishery, while above this stock, harvests
greater than \(H_{MSY}\) will be needed to bring the stock back to
\(B_{MSY}\). (This intuition must be adjusted slightly in the case of
economic discounting of future profits, but is otherwise quite general,
see Clark (1990)). This strategy is known as constant escapement, since
a constant sock size \(B_{MSY}\) escapes harvest each year. While both
converge to the same long-term biomass and long-term yield, these
differences away from equilibrium between the mortality-based approach
that dominates fisheries management practice and the escapement-based
approach that became the focus of theory will result in different
behaviors under uncertainty. Intuitively, any uncertainty, either in the
population growth model (stochasticity), or measurement error, should
seem to justify harvesting fewer fish, as is built into the TAC limits
discussed above. Yet formal treatments of these issues have so far
required approximations, and these approximations have driven
paradoxical conclusions which we briefly summarize here.

\subsection{\texorpdfstring{Reed's Paradox:
\(S = D\)}{Reed's Paradox: S = D}}\label{reeds-paradox-s-d}

The first of these we shall refer to as Reed's Paradox for stochastic
growth, owing to the mathematical proof provided in Reed (1979) which
demonstrates that, under sufficiently general assumptions, the optimal
escapement \(S\) for a population under stochastic growth, is identical
to the optimal escapement \(D\) of a deterministic population,
\(S = D = B_{MSY}\). This surprising result suggests that in going from
a world where a manager has perfect knowledge of absolutely everything
into a scenario where the manager faces considerable uncertainty about
the future state of the world, no additional precaution is needed. This
also provides the manager with a remarkably convenient mechanism for
determining the optimal harvest policy: rather than rely on
computationally intensive Stochastic Dynamic Programming to determine
the optimal policy, a simple derivative will suffice, which is all that
is needed to solve the maximum of the deterministic problem.

\subsection{Clark's Paradox}\label{clarks-paradox}

@@

Unfortunately, Clark and Kirkwood (1986) was unable to solve the
resulting problem exactly, but had to adopt an almost equally
troublesome assumption:

@@

Unfortunately for Clark and Kirkwood (1986), this ``simplifying
assumption'' serves to squash most of the measurement error, and their
results instead only deepened the paradox, finding policies that become
even \emph{less} cautious as uncertainty increases:

@@

Despite these notes of caution, both Clark and Kirkwood (1986) and Sethi
et al. (2005) ultimately attempt to rationalize this counter-intuitive
conclusion rather than reject it. Others have also attempted to
introduce measurement error, but always relying on various
approximations that similarly alter the problem. The most common such
assumption is to assume from the outset that the solution must be of
`constant-escapement' type (Ludwig and Walters 1981; Roughgarden and
Smith 1996; Engen et al. 1997; Moxnes 2003). Here we will illustrate the
use of powerful modern algorithms that will permit a more direct
solution to the case of measurement uncertainty. In addition to
resolving this paradox, we will demonstrate this direct solution also
out-performs current more heuristic rules of thumb such as those
currently used in fisheries management to account for uncertainty.

\subsection{POMDPs: An optimal treatment of measurement
uncertainty}\label{pomdps-an-optimal-treatment-of-measurement-uncertainty}

As discussed above, absent the short-cut provided by Reed's theorems --
solving for the optimal management policy in a stochastically
fluctuating population -- involves a Markov Decision Process (MDP),
commonly referred to in ecological literature by its solution method,
Stochastic Dynamic Programming (SDP). The constant escapement (CE)
strategy considered here is the result of such an MDP or SDP analysis.
Once measurement uncertainty is introduced, the problem becomes a
Partially Observed Markov Decision Process (POMDP, not be confused with
a partially observed or hidden Markov model, HMM, in which no decision
process is involved). The POMDP problem is posed identically to that of
the MDP problem, with the addition of an observation process, but cannot
be solved using SDP directly. The POMDP problem for fisheries question
considered here can be summarized as follows:

\begin{itemize}
\item
  Transition function (state equation): \(T(x_t, x_{t+1}, a_t)\): the
  probability that a system is in state \(x_{t+1}\) at time \(t+1\)
  given that it began in state \(x_t\) at time \(t\) and the manager
  took action \(a_t\). In our context, this relationship is given by the
  Gordon-Schaefer stock recruitment function \(f\) with normally
  distributed growth uncertainty
  \(x_{t+1} \sim \mathcal{N}(f(x_t,a_t), \sigma_g)\), truncated at zero
  to exclude negative population sizes.
\item
  Observation function: \(O(x_t,y_t,a_{t-1})\) the probability of
  observing state \(y_t\) given a system in state \(x_t\). In principle,
  the action chosen can influence the precision of the observation. In
  our case, we simply assume normally distributed errors around the true
  state, \(y_t \sim \mathcal{N}(x_t, \sigma_m)\), truncated at zero to
  exclude negative population sizes.
\item
  Utility function: \(U(x_t,a_t)\), the utility received at time \(t\)
  for taking action \(a_t\), given that the system is in state \(x_t\).
  For simplicity of analysis, we will simply set the utility to be equal
  to the harvested stock: \(U(x_t, a_t) = \min(x_t, a_t)\), indicating
  that realized harvest cannot be negative. This choice ensures that in
  the case of no uncertainty (\(\sigma_g = \sigma_m = 0\)), the optimal
  solution matches that expected under a simple MSY calculation. More
  realistic utility functions may include diminishing returns with
  increasing harvest (supply and demand effects), and the cost of
  fishing, both of which act to suppress large harvests. By focusing on
  a simple utility we can be sure that our comparison to MSY is driven
  by the treatment of uncertainty rather than merely differing economic
  assumptions.
\end{itemize}

The optimization problem is to select the action \(a_t\) that will
maximize the net present utility over all time. Future utility may be
discounted by a factor \(\gamma\), so that a value \(V\) in \(t\) year
is valued at \(\gamma^t V\) today. Numerically, each of these functions
are defined over a discrete set of possible states, observations, and
actions, and can thus be represented as a collection of matrices or
tensors. While the solution method for MDP problems, stochastic dynamic
programming (SDP) is well known and readily implemented in ecological
problems (e.g Mangel and Clark 1988; Marescot et al. 2013), the POMDP
problem cannot be efficiently solved using SDP. While not unknown to the
conservation literature (Williams 2011), algorithms for POMDP have
historically scaled quite poorly, and their application has been
restricted to contexts with only a handful of possible states and
actions (e.g. Chadès et al. 2008, 2011; Fackler and Haight 2014; Fackler
and Pacifici 2014). POMDP algorithms have a substantial literature
starting almost contemporaneously to the work of Reed (1979) (see
foundational papers by Smallwood and Sondik (1973) and Sondik (1978))
and remain an active area in artificial intelligence (e.g. Kaelbling et
al. 1998; Pineau et al. 2003). By adapting recent algorithmic
developments (Kurniawati et al. 2008) from that field we are able to
find solutions for the considerably more complex problems such as the
case of fisheries management, which requires on the order of 100 states
and actions to provide sufficient numerical resolution. As the details
for solving such problems numerically are already well documented we
will not rehash them here, but instead provide a preliminary R package
(Boettiger et al. 2018) which provides an implementation of the SARSOP
algorithm and the fisheries POMDP problem, which can be used to
replicate any of the results presented here and explore other
variations. Details of the analysis presented here including annotated
code to reproduce and further explore all of the following results is
provided in the supplementary material.

\section{Results}\label{results}

>>>>

  \begin{abstract}
  Ecological management and decision-making typically focus on uncertainty
  about the future, but surprisingly little is known about how to account
  for uncertainty of the present: that is, the realities of having only
  partial or imperfect measurements. Our two primary paradigms for
  handling decisions under uncertainty -- the precautionary principle and
  optimal control -- have so far given entirely contradictory results.
  This paradox is best illustrated in the example of fisheries management,
  where many ideas that guide thinking about ecological decision making
  were first developed. We find that simplistic optimal control approaches
  have repeatedly concluded that a manager should increase catch quotas
  when faced with greater uncertainty about the fish biomass. Current best
  practices take a more precautionary approach, decreasing catch quotas by
  a fixed amount to account for uncertainty. Using comparisons to both
  simulated and historical catch data, we find that neither approach is
  sufficient to avoid stock collapses under moderate observational
  uncertainty. Using partially observed Markov decision process (POMDP)
  methods, we demonstrate how this paradox arises from flaws in the
  standard theory, which contributes to over-exploitation of fisheries and
  increased probability of economic and ecological collapse. In contrast,
  we find POMDP-based management avoids such over-exploitation while also
  generating higher economic value. These results have significant
  implications for how we handle uncertainty in both fisheries and
  ecological management more generally.
  \end{abstract}
   \begin{keyword} Measurement uncertainty POMDP Decision theory Fisheries Conservation\end{keyword}
 \end{frontmatter}

Imperfect information is ubiquitous in ecological management and
conservation decision making. While the pressing concerns of global
change have put the spotlight on \emph{forecasting}, i.e.~the
uncertainty of the future (e.g. Petchey et al. 2015), management
decisions must also contend with uncertainty of the \emph{present}: How
many fish are in the sea today? What regions harbor the most fragile
biodiversity? How have invasive species transformed native communities?
A recognition of the importance of uncertainty has become widespread in
the context of estimating ecological models from available data, thanks
in large part to the rise of approaches such as Hierarchical Bayesian
Modeling, along with convenient software tools for implementing these
methods. In sharp contrast to that shift, the treatment of uncertainty
in decisions -- how the uncertainty in models should be reflected in
scientific recommendations, remains far less developed, more
qualitative, and under-served by available tools. We illustrate how
uncertainty in measurement has a dramatic negative impact on outcomes
under a range of existing approaches, which can be avoided through the
use of more formal approaches that have since been refined in fields
such as robotics. This insight both resolves a long-standing paradox in
fisheries management and has important implications for how ecosystem
management deals with uncertainty more generally.

Approaches for decision making under uncertainty in ecological systems
can be divided into two camps: (1) those based upon ``optimal control''
solutions, and usually favored by natural resource economists, and (2)
those based in heuristic methods such as scenario planning, resilience
thinking, and precautionary rules of thumb, more commonly found in both
ecological literature and actual practice (Fischer et al. 2009; Polasky
et al. 2011). While researchers have recognized the need to unify the
transparent and quantitative algorithmic approach of optimal control
with the greater complexity and uncertainty of real ecosystems that is
acknowledged by heuristic methods, computational barriers to doing so
have stymied this progress. We illustrate how the limitations of these
approaches have manifested in the example of fisheries management, and
present a new approach that can combine the reality of measurement
uncertainty and the rigor of optimization to resolve a long-standing
paradox and suggest a more robust approach to management. We compare the
results of managing simulated fish stocks under (A) the existing optimal
control approach, (B) a heuristic approach that is more typical of
today's management practices, and (C) a newly proposed approach that
provides a more rigorous treatment of uncertainty.

The challenge of making decisions under uncertainty is obviously not
unique to fisheries, and the existing methods we critique here can be
found in a wide range of population and ecosystem management challenges,
including fire ecology (Richards et al. 1999), invasive species
({\textbf{???}}; Blackwood et al. 2010), disease outbreaks (Shea et al.
2014; Li et al. 2017), and protecting biodiversity (Dee et al. 2017;
Iacona et al. 2017). Yet fisheries conservation and management has long
been both a crucible and proving ground for the theory of ecological
management more generally, including topics such as adaptive management
(Walters and Hilborn 1978), ecosystem-based management (Levin and
Lubchenco 2008) and resilience thinking (Holling 1973; May 1977), while
also giving rise to the sub-discipline of resource economics (Gordon and
Press 1954; Schaefer 1954; Beverton and Holt 1957), thanks to its global
relevance, long history, and readily available data. Despite the
generality of the models and issue of measurement uncertainty considered
here, any application must be grounded in both the history and context
of the particular decision problem. In light of this, we begin with a
brief background on existing decision theory approaches common in
fisheries management.

\hypertarget{the-decision-problem}{%
\section{The Decision Problem}\label{the-decision-problem}}

We consider the management problem of setting catch quotas for a marine
fishery in the face of imperfect information about the current stock
size and uncertainty about future recruitment. We seek to determine the
sequence of actions \(h_t\) for \(t \in [1, ... \infty]\) that maximize
the net present value (discounted sum of all future profits) of the
fishery. We will denote the discount rate \(\gamma\). For simplicity we
will once again follow classic theory and assume a fixed price for fish,
(equivalently, measuring our value in units of discounted fish rather
than discounted dollars), though again this assumption is easy to relax.
Each year, the manager also obtains an estimate \(y_t\) of the true
\(x_t\) stock size subject to some measurement uncertainty \(\xi_t\)

\begin{equation}
y_t = \xi_t x_t 
\end{equation}

@@

\begin{equation}
\max_{ \lbrace h_t \rbrace } \mathbb{E} \left \{ \sum_{t=0}^{\infty} \gamma^t \cdot h_t \right \}
\end{equation}

@@

\hypertarget{population-model}{%
\subsection{Population model}\label{population-model}}

@@

\begin{equation}
x_{t+1} = \varepsilon_t (x_t-h_t) r  \left(1 - \frac{(x_t-h_t)}{K}\right) \label{Gordon-Schaefer}
\end{equation}

where \(x_t\) is the current stock size, \(h_t\) the harvest chosen that
year, with parameters \(r\) giving the individual growth rate, and \(K\)
the carrying capacity and \(\varepsilon\) representing stochastic
recruitment in a variable environment. We have written this in terms of
\(x_t - h_t\) to underscore that we assume the convention of observe,
harvest, recruit. We will assume for simplicity
\(\varepsilon \sim \mathcal{N}(1, \sigma_g)\). This and similar models
are widely used in large scale analyses across diverse stocks (e.g.
(Costello et al. 2016; Britten et al. 2017)), and form the basis for
much of bioeconomic theory (Clark 1990). As such, it will be easier to
compare our results against intuition and classic theory, and avoid the
possibility of differences arising only because of some particular
subtle assumption hidden in a more complex model.

It is important to remember that both existing methods we discuss and
the POMDP approach proposed here can or are already applied to more
complex models as well, including those with age or stage structure.
Likewise, while we consider the model and its parameters as ``given,''
it should be understood to mean that this in practice comes as the
result of some model choice and statistical estimation from historical
data, frequently accounting for uncertainty in both the process and the
measurements (e.g. Dichmont et al. 2016). As we shall see here, having
included uncertainty in the model \emph{estimation} in no way excuses us
from having to also deal once again with that uncertainty in the
decision process.

\hypertarget{current-theory-and-practice}{%
\section{Current Theory and
Practice}\label{current-theory-and-practice}}

@@

\hypertarget{uncertainty-in-practice-msy-tac}{%
\subsection{Uncertainty in Practice: MSY \&
TAC}\label{uncertainty-in-practice-msy-tac}}

Maximum Sustainable Yield (MSY) remains the basis of international law
(including the UN, IWC, IATTC, ICCAT, ICNAF; Mace (2001)) and a familiar
standard of management in terrestrial ecosystems as well as aquatic
(Clark 1990). Critics have for some time observed the limitations of
harvesting at MSY in face of uncertainty in stock sizes and population
dynamics (Larkin 1977; Botsford et al. 1997). Many US fisheries reflect
this uncertainty through a series of adjustments that effectively reduce
the target fishing mortality level to reflect this uncertainty.
Typically, a stock assessment model provides a best estimate of
\(B_{MSY}\) and corresponding mortality \(F_{MSY}\) is used to define
the stock Over-Fishing Limit, (OFL). Based on this, a somewhat lower
level is set as the Allowable Biological Catch (ABC), reflecting
uncertainty in the stock assessment. To reflect possible uncertainty
between reported and actual catch, the ABC is reduced somewhat further
to define the Total Allowable Catch, (TAC), which forms the basic unit
of management for many such fisheries. To reflect this process, our
analysis will also consider policies in which the harvest quota is set
at 80\% of the level expected under the MSY policy:
\(H_{TAC}(t) = 0.8 \cdot F_{MSY} \cdot B_t\), for a biomass estimated at
\(B(t)\) in year \(t\). To distinguish this approach from MSY, we will
refer to this approach as a TAC policy. This more closely represents a
heuristic (e.g. Hilborn 2010) or resilience-based approach (sensu
Fischer et al. 2009) than an optimization-based policy. Importantly,
this approach shares the fundamentally stationary assumptions of an MSY
policy by defining a constant mortality rather than a dynamic policy.
Thus, despite being more cautious overall and generating lower economic
yield than a constant escapement policy, TAC policies continue to
harvest at non-zero rates even if a stock falls below \(B_{MSY}\), while
the constant escapement policy does not.

It is important to note that actual management strategies are quite
diverse. Our focus on comparisons TAC, MSY, and Constant Escapement (CE,
also refereed to as MEY, see below) reflects only the subset of current
practices. Another increasingly common approach to address uncertainty
in both process and models is known as Management Strategy Evaluation
(MSE, Smith 1994; {\textbf{???}}; Bunnefeld et al. 2011; Punt et al.
2016). MSE performs forward simulations under a suite of pre-selected
candidate strategies to determine which strategy achieves the best
expected outcomes. Crucially, by limiting the analysis to a suite of
candidate strategies over a finite number of replicate simulations, this
approach is able to accommodate substantially greater complexity and
uncertainty than has been possible with more formal optimal control
approaches we discuss below. Optimal control solutions rely on dynamic
programming to consider all possible strategies (all possible sequences
of harvest under all possible realizations of stock dynamics). As such,
MSE can be viewed as an approximation to the optimal control solutions
considered below, a notion that has been formalized in the literature on
approximate dynamic programming (Nicol and Chadès 2011).

\hypertarget{optimal-management}{%
\subsection{Optimal management}\label{optimal-management}}

So far, attempts to provide a more formal basis for managing uncertainty
than the heuristic adjustment of catch limits described above have
largely foundered in a series of paradoxes. Without uncertainty, the
theoretical policies are quite intuitive: The result derived by Clark
(1973) for an optimal dynamic strategy to replace the equilibrium
solution of MSY can be summarized as: If the stock is most productive at
\(B_{MSY}\), then obtain \(B_{MSY}\) as quickly as possible. Thus, at
any level below \(B_{MSY}\), the economically optimal thing to do is to
completely shut down the fishery, while above this stock, harvests
greater than \(H_{MSY}\) will be needed to bring the stock back to
\(B_{MSY}\). (This intuition must be adjusted slightly in the case of
economic discounting of future profits, but is otherwise quite general,
see Clark (1990)). This strategy is known as \emph{constant escapement}
(CE), since a constant sock size \(B_{MSY}\) escapes harvest each year.
Because CE and MSY converge to the \emph{same long-term biomass and same
long-term average yield}, CE can appear to be merely a more formal
justification for MSY. However, in practice the two approaches are very
different: because natural populations fluctuate year-to-year, under CE,
a different harvest quota must be set each year to achieve the constant
escapement. Because CE is derived as the economic optimum (assuming
perfect measurements), it is sometimes referred to as Maximum Economic
Yield, (MEY; Burgess et al. 2018), or as Rights Based Fisheries
Management (RBFM, Costello et al. 2016).

In the face of a naturally fluctuating population, previous arguments
made by Gordon and Press (1954); Schaefer (1954) or Clark (1973) no
longer hold. The optimal harvest of a population under a model such as
the stochastic Gordon Schaefer, Eq \eqref{Gordon-Schaefer}, is an
example of a Markov Decision Problem (MDP). This is not to be confused
with the the problem of \emph{estimating} parameters or posteriors of a
Markov process model, e.g.~through typical Bayesian analysis -- recall
this is a decision problem, in which the model has already been
estimated and now we are tasked with choosing a sequence of future
harvests based on that model. Such MDP problems are commonly referred to
in ecological literature by its solution method, Stochastic Dynamic
Programming (SDP) (Marescot et al. 2013), widely used in both ecological
management and animal behavior (Mangel 1985; Mangel and Clark 1988).
However, thanks to a mathematical proof provided by Reed (1979),
fisheries managers have largely been able to bi-pass the need for SDP.
But this convenient result also opened the door to paradox of
uncertainty that would persist for the next four decades.

\hypertarget{reeds-paradox-s-d}{%
\subsection{\texorpdfstring{Reed's Paradox:
\(S = D\)}{Reed's Paradox: S = D}}\label{reeds-paradox-s-d}}

Reed (1979) proved that, under sufficiently general assumptions, the
optimal escapement \(S\) for a population under stochastic growth, is
identical to Clark (1973)'s optimal escapement \(D\) for a deterministic
population, \(S = D = B_{MSY}\). This surprising result suggests that in
going from a world where a manager has perfect knowledge of absolutely
everything into a scenario where the manager faces considerable
uncertainty about the future state of the world, no additional
precaution is needed.

\hypertarget{clarks-paradox}{%
\subsection{Clark's Paradox}\label{clarks-paradox}}

@@

Clark and Kirkwood (1986) were unable to solve the resulting problem
exactly, but had to adopt an almost equally troublesome assumption:

@@

Unfortunately, this ``simplifying assumption'' serves to squash most of
the measurement error, and their results instead only deepened the
paradox, finding policies that become even \emph{less} cautious as
uncertainty increases:

@@

Despite these notes of caution, both Clark and Kirkwood (1986) and Sethi
et al. (2005) ultimately attempt to rationalize this counter-intuitive
conclusion rather than reject it. Intuitively, uncertainty in the
population growth model (stochasticity) or measurement error, would seem
to justify harvesting fewer fish. Instead, existing optimal control
theory has presented only the paradoxical results that uncertainty
either should not matter at all, or that we should increase rather than
decrease harvests in response to increased uncertainty. Numerous other
attempts have been made to resolve this paradox through alternative
assumptions, the most common being to assume from the outset to look
only for `constant-escapement' type policies (Ludwig and Walters 1981;
Roughgarden and Smith 1996; Engen et al. 1997; Moxnes 2003). Owing to
the ``mathematical difficulty'' Clark first observed, none have
attempted a direct solution which we will present here.

In practice, TAC-managed fisheries are more common that CE-managed
fisheries, at least in the United States. CE-based regulation is limited
primarily to salmon, species for which population estimates may be more
precise and thus closer to Reed's assumption. We are not aware of any
fishery that has put Clark's Paradox into practice, harvesting more in
response to increased uncertainty. As we shall illustrate, this is
particularly fortunate; though approaches that properly account for both
forms of uncertainty can perform even better.

\hypertarget{pomdps-an-optimal-treatment-of-measurement-uncertainty}{%
\subsection{POMDPs: An optimal treatment of measurement
uncertainty}\label{pomdps-an-optimal-treatment-of-measurement-uncertainty}}

Why does the introduction of measurement uncertainty make the decision
problem so much harder? Consider the thought experiment of managing a
fish stock where in each of the previous four years stock assessments
have put the population size at, say, 51, 54, 49, 51, (say, measured in
percent of carrying capacity, \(K\)). Under a constant escapement target
of 50, the harvest quota would be 1, 4, 0, 1, respectively. Note that
this control rule depends only on the measurement for the year in
question, which is assumed to be measured without error. If we then
measure a stock size of 75, an assumption that this measurement is made
without error would justify the sudden increase to a quota of 25. Yet if
we admit the possibility of measurement error, this sudden bump starts
to look very suspicious in light of all the previous observations.
Perfect measurements let us exploit the Markov property of the model --
the most recent measurement tells us all we need to know. In contrast,
the introduction of measurement uncertainty breaks this Markovian
assumption: a rational decision maker would suspect this 75 to be an
over-estimate based on prior observations. Under imperfect observations,
we no longer have a Markov Decision Process (MDP) over the observed
state space. Rather, we have a Partially Observed Markov Decision
Process (POMDP). A POMDP policy cannot be expressed as a policy function
at all (much less as policy function with a single constant-escapement
target, as much previous work has assumed). Instead, a POMDP policy can
be thought of as depending not only on the most recent observation, but
the recent observation combined with a \emph{prior belief} about what
state the system is in, which itself is determined by all past
observations. In this manner, the POMDP can consider how much weight to
give the most recent observation based on how it compares to prior
beliefs.

Algorithmic innovations from the robotics community have now made it
possible to numerically solve non-trivial POMDP problems. The need to
consider all past observations becomes particularly demanding given the
forward-looking nature of a decision problem. Not only must we consider
all possible sequences of future actions over all possible sequences of
future states, but we must add to that all possible sequences of future
measurements. The computational complexity involved has historically
limited POMDP applications to contexts with only a handful of possible
states and actions (e.g. Chadès et al. 2008, 2011; Fackler and Haight
2014; Fackler and Pacifici 2014), insufficient to capture fisheries
models appropriately. Meanwhile, methods to solve decision problems with
imperfect information have advanced steadily in the field of robotics
and artificial intelligence. Almost contemporaneous with Reed's work,
papers by Smallwood and Sondik (1973) and Sondik (1978) laid
mathematical foundations for algorithms that could efficiently solve
POMDP problems (e.g. Kaelbling et al. 1998; Pineau et al. 2003). More
recently, driven by demands in areas such as autonomous vehicle decision
making, newer point-based approximation algorithms such as SARSOP
(Kurniawati et al. 2008) have made it possible to solve POMDPs with 100s
of states and actions. We have adapt this algorithm for application to
the fisheries context, and provide an efficient implementation as a
preliminary R package (Boettiger et al. 2018) which can be used to
replicate and further explore results presented here. Details of the
analysis presented here including annotated code to reproduce and
further explore all of the following results is provided in the
supplementary material.

Precise definitions of both the (fully observed) MDPs and POMDPs (not to
be confused with Markov Process Models and Hidden Markov Process Models
-- methods for which estimate the model but do not involve the feedback
of any decision process) in an ecological context can be found in
Williams (2011), or in classic work from the engineering and robotics
community (e.g. Smallwood and Sondik 1973, and @Sondik1978). For
convenience, we summarize the POMDP definition and common notation here.
The POMDP problem is posed identically to that of the MDP problem, with
the addition of an observation process The POMDP problem for fisheries
question considered here can be summarized as follows:
